Unveiling Robustness Threats by Mining the Exception Stack Traces of over 600 Android Projects

### 2014-09-06 android-stacks.tex
all over

Until recent years, the availability of data, reporting such failures, was scarce. 
It comprised only few large scale projects (e.g. the Mozilla crash report dataset\footnote{https://code.google.com/p/promisedata/issues/detail?id=88}).
Recently we have observed a rapid increase in the number of open
source projects that make their version control systems as well as the issue tracking systems
publicly available in repository hosting sites such as GitHub and Google Code.
Developers and users of open-source projects uses such provided
 infra-strucute to share information (e.g., exception stack traces~\cite{bettenburg2008makes,schroter2010stack}) across 
the software development.

@@

This study performs a post mortem analysis of the exception stack traces embedded on issues
reported on Android projects hosted in GitHub and Google Code. 
The goal of this study was to: investigate whether the reported exception stack traces
 can reveal common fragilities on the exception-related code.  
We call ``fragilities" characteristics on the exception-related code 
that favor the introduction of failures such as uncaught 
exceptions -  contributing to decrease the applicaiton robustness.
To guide this investigation we compiled general guidelines on how to use
exceptions proposed by Gosling~\cite{gosling2000java},
Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective}.
Then, using a custom tool called ExceptionMiner,
 which we develop specifically for this study, we mine stack traces from the issues reported 
on 482 Android projects hosted in GitHub and 157 projects hosted in Google Code.
Overall 159,048 issues were analyzed and 6,005 stack traces were extracted from them.
The exception stack trace analysis was augmented with additional information
extracted from JDK, Android Plaform and the applications using byte code 
and source code analysis. Manual inspection was used to leverage the understanding 
of exception stack traces and support further discussions and insights.

@@

The Android Application Framework supports an event-driven development model. 
Among the basic elements that comprise an Android application, we can cite the
 Activity class. The Activity is the main class of an Android app (there is usualy one
Activity per application window). Each Activity defines a set of hook methods  
(e.g., OnCreate(), onPause()) that are dispatched automatically by the Android 
framework (by inversion of control), in response to infrastructure and application events.
The Android Application Framework also provides several supporting APIs, such as 
the APIs to support concurrent development (e.g. AsyncTask). 
Such APIs particularly useful due to the high demand for responsiveness 
in mobile environment (mobile apps can easily become 
unresponsive due to resource limitation and excessive network access).

@@

By convention an Error represents an unrecoverable condition which usually results
from failures detected by the Java Virtual Machine, such as OutOfMemoryError and
normally should not be handled inside the application. The user-defined exceptions 
can be either checked or runtime. The Java specification~\cite{gosling2000java} 
suggests that user-defined exceptions should be checked (because doing so 
the callers of a method will know about the exceptions that a method can throw and so 
they can decide what to do about them). However, runtime exceptions can also be 
used to represent user-defined exceptional conditions.In fact, there is a long-lasting debate 
about the pros and cons of both approaches~\cite{javatut,stackoverlow,debate}
%\footnote{152 questions in Stackoverlow are related to this debate}.

@@

%In Java an exception can be thrown in one of the following
%circumstances~\cite{gosling2000java}: 
%(i) explicitly thrown when a throw statement is executed; 
%(ii) implicitly thrown by the JVM when the evaluation of an expression
% violates the normal semantics of language, also referred as a coding error
% (e.g., out-of-bounds array index, division-by-zero, access to a null reference); or 
%(iii) implicitly thrown by the JVM due to an internal error or resource limitation (e.g.,
%OutofMemoryError). 
%\todo{This sub-section is quite similar to the section introduction, shouldn't
%we merge them?}

%When an exception is thrown, it causes a transfer of control from the point
%where the exception occurred to a point that can be specified by the programmer
%(exception handler) - in Java it is represented by the try-catch block .

@@

%<AVALIAR MOVER ESTE STACK TRACES PARA UMA FIGURA QUE MOSTR COMO ELE É QUEBRADO...
%PELO EXCEPTION MINER E AS INFORMAÇÕES VAO PARA O BANCO...>

% \subsection{Public Project Repositories and Issue Trackers} 
% ONE OF THE REVIEWERS COMPLAINED ABOUT THE ABSENCE OF EXPLANATIONS OF WHAT AN ISSU IS...
% As mentioned before, currently more and more projects are becoming available in 
% public project repositories such as GitHub and Google Code.
% Such repositories provides facilities such as....
% The issue trackers available in such projects ....

@@

\emph{IV- All exceptions explicitly thrown by a library/framework reusable code should be documented.}
(\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}).
The exceptions thrown by a method are an important part of methods interface,
and is required to use the method properly. The checked exceptions are already
part of the  methods signature, and the method caller is aware of the checked
exceptions being thrown by it. According to~\cite{bloch2008effective}, it is
also wise to document the explicitly thrown runtime exceptions\footnote{This
excludes the implicitly signaled runtime exceptions related to programming
mistakes} as carefully as checked exceptions, especially when developing
 library/framework reusable code. Doing so, the clients of such ibrary/framework
will be aware of all the exceptions a method can throw and can design the code
to deal with them ~\cite{Robil00}.

If the developer fails to do follow such practices (specially when developing 
utility code) it will be difficult or even impossible for the caller to 
make effective use of such method~\cite{wirfs2006toward, bloch2008effective},
and protect the system against unforeseen exceptions thrown by the utility code,
which can lead to the uncaught exceptions - one of the main causes of 
systems crashes. These best practices guided our exploratory study dscribed next.

%Next section discusses the
%study procedure and the practices' misuses that emerged from this study.

%This study focus on the analysis of crash information (related to uncaught exceptions) of a family of Android applications
%aiming at  extracting their common characteristics and looking for similar vulnerabilities on applications and underlaying platform.
%In Java when an application crashes due to an uncaught exception ... 
%Exception Stach trace is usually (procurar por artigo).

%To answer this question we selected 
%The goal of this exploratory mining study was to extract exception stack traces embedded on issues of Android projects 
%(hosted in GitHub and Google Code), looking for its \emph{common characteristics} and investigating whether
% they could reveal \emph{common fragilities} on the exception code of both the applications and the underlaying  framework. 
%Our study focused on open-source apps, since the information needed to perform our study
%cannot be retrieved from commercial apps, whose issue report systems and 
%source codes are not publicly available. The Android open-source apps were also 
%target of other research studies ~\cite{Linar13,ahimed}.

%extract exception stack traces embedded on issues of Android projects 
%(hosted in GitHub and Google Code), looking for its \emph{common characteristics} and investigating whether
% they could reveal \emph{common fragilities} on the exception code of both the applications and the underlaying  framework. 
%Our study focused on open-source apps, since the information needed to perform our study
%cannot be retrieved from commercial apps, whose issue report systems and 
%source codes are not publicly available. The Android open-source apps were also 
%target of other research studies ~\cite{Linar13,ahimed}.

@@

The goal of this exploratory mining study was to investigate whether the exception stack traces 
embedded on issues of Android projects (hosted in GitHub and Google Code), can reveal 
\emph{common fragilities} on the exception code of both the applications and the underlaying  framework. 
Our study focused on open-source apps, since the information needed to perform our study
cannot be retrieved from commercial apps, whose issue report systems and 
source codes are not publicly available. The Android open-source apps were also 
target of other research studies ~\cite{Linar13,ahimed}.

@@

To support this investigation, we developed a tool called ExceptionMiner (see Section ~\ref{sec:exceptionminer})
which extracts the exception stack traces embeeded on issues, 
and combines stack trace information with source code and bytecode 
analysis. Such combination was needed to enable the identification of 
fragilities on the exception-related code. Moreover, manual inspections were also used to leverage
 the understanding of stack traces and support further discussions and insights.
In this study we explore the domain quantitatively and highlight interesting cases by 
exploring cases qualitatively.  Figure~ref{fig:overview} gives an overview of the study, whose tasks are detailed next.

@@

This study used the dataset provided by the GHTorrent project~\cite{Gousi13}, 
an off-line mirror of the data  offered through the Github API.  
To identify Android projects, we performed a case insensitive search for the
term \textsf{android} in the repository's names and short descriptions.  
Up to 23 Feb 2014,  when we queried GHTorrent, such heuristic could
 find 2.542 repositories. Running the ExceptionMiner tool (described next)
 in this set we observed that 589 apps had at least one issue containing a stack trace.

Then we performed further cleanup, inspecting the site of every Android
reporting at least one stack trace, to make sure that they represented real
mobile apps. During this clean up 106 apps were removed because they were either
example projects (i.e., toy projects) or tools to support Android development
(e.g. Selendroid, Roboeletric - tools to support the testing of Android apps).
The filtered set consisted of 482 apps. This set of 482 projects contained overall 31,592 issues from which 4,042 exception stack traces 
were extracted. Issues on Github are different from issues on dedicated bug tracking tools such as 
Bugzilla and Jira. The most important difference is that there are no predefined fields
  (e.g. severity and priority). Instead, Github uses a more open ended tagging system, on which
repositories are offered a pre-defined set of labels, but repository owners can modify 
them at will. Therefore, an issue may have none ore an arbitrary set of labels depending 
on which repository it was created. Table ~\ref{tab:lables} illustrates the ocurrences of different lables 
on the issues including exception stack traces.

@@

\begin{table}
  \centering
  \begin{tabular}{lr|lr}
    \hline
     \multicolumn{2}{c}{\bfseries{GitHub}} &  \multicolumn{2}{c}{\bfseries{Google Code}} \\
      \bfseries{Lable} &  \bfseries{\% Occurrences} &  \bfseries{Lable} &  \bfseries{\% Occurrences} \\
    \hline
empty &	54,24\% & Defect &	91,96\% \\
Defect &	39,56\%  & Enhancement  &	3,16\% \\
Enhancement &	0,57\% & Task	& 1,37\% \\
Support &	0,52\% & empty &	1,12\% \\
Problem &	0,36\% & StackTrace &	0,70\% \\
Others &	4,74\% &  Others &	1,68\% \\   
  \hline
  \end{tabular}
  \caption{Lables on issues including exception stack traces.}
  \label{tab:lables}
\end{table}

Based on the assumption that regardless the issue type, every exception stack
trace contains relevant information concerning the exception structure of the
projects analyzed - and therefore can review fragilities on the exception-related code -  
we opted for not restricting the analysis only on defect
issues\footnote{We conducted the same analysis on the defect issues and the top
exceptions are similar to the ones mentioned on defect issues. Due to space
limitation we limit to present the general analysis here. More detailed analysis
on the defect issues can be found at:
\url{https://github.com/souzacoelho/exceptionminer}}.
 Moreover, issues and pull requests are dual on Github; all pull requests have a corresponding 
``backing'' issue which are automatically generated. Therefore, we excluded such automatically generated
issues from our analysis.

\subsubsection{Android Apps in Google Code}
Google Code contains largely used open-source Android apps (e.g. K9Mail \footnote{K9Mail moved to Github but as a way of not loosing the project history 
it advices their users to report bugs on Google Code issue tracker.}), however, differently from GitHub, Google Code does not provide an API to access the information related
 to hosted projects \footnote{Google code used to provided a Web service to such repository, but it was deactivated in June 2013 what Google called a "clean-up action".}.
To overcome this limitation we needed to implement a Web Crawler (incorporated in ExceptionMiner tool described next) that navigated 
through the web interface of Google Code projects extracting all issues and issue comments and storing in a relational database for later analysis.

To identify Android projects in Google Code, we performed a similar heuristic: we performed a case insensitive search 
(on Google Code search interface) for the term ``android". On January 2014, when we queried Google Code, such heuristic could
 returned a list of 788  projects. This list comprised the seeds sent to our Crawler.

From this set, 724 defined at least 1 issue. Running the ExceptionMiner tool (described next)
 in this set we observed that 183 apps had at least one issue containing an exception stack trace.
 Then we performed further cleanup (similar to the one described previously) inspecting the site 
of each project. As a result we could identify 157 Android projects.  This set contained overall 127,456 issues,
 from which 1,963 exception stack traces were extracted. Table ~\ref{tab:lables} illustrates the ocurrences of different lables 
on the issues including exception stack traces. Differently from GitHib, in Google Code most of 
the issues were labled as ``Defect". However, based on the same assumption described for the GitHub repository
 we considered all issues reporting stacks (regardless its lables).

@@

The ExceptionMiner is modular mining tool which is able to connect to issue repositories, 
extract issues, mine stack traces from them, distill exception stack trace information,
and enable the execution of different analyses by combining exception stack 
trace information with bytecode and sourcecode analysis. The main components of ExceptionMiner are as follows:

\emph{\textbf{Repository Connectors.}} This component enables the connection 
with issue repositories. In this study two main connectors were created: one which connects to 
GHTorrent database, and Google Code connector which is comprised of a Web Crawler,
 that can traverse Google Code Web interface and extract the project issues. 
Projects matadata and the issues associated to them are stored on a relational 
database by this components.

\emph{\textbf{Exception Stack Trace Distiller.}}
This component distills the information that composes a stack trace.
 Some of the attributes extracted from the stack trace are:
 the root exception and its signaler, the exception wrappers and their corresponding signalers. 
This component also distills fine grained information of each attribute such as: the classes and packages associated to them.
This component is based on a combination between a regular expression based parser 
with heuristics able to identify and filter exception names and stack traces inline with text. 
In contrast to existing issue parsing solutions such as Infozilla, the parser
created in this work: (i) can discover stack traces inlined with logs files \footnote{In several 
exception stack traces, the exception frames were preceeded by logging information e.g., 
"03-01 15:55:01.609 (7924): at android.app.ActivityThread.access\$600(ActivityThread.java:127) 
"which could not be detected by existing tools.)}.

@@

As mentioned before, the tool enables a set of exception stack traces analyses by combining exception stack 
trace information with bytecode and sourcecode analysis. Each analysis is implemented as a different module 
on ExceptionMiner. Next, we present core analyses which were implemente on ExceptionMiner. Such analyses 
provide aditional information to be used on further analyses described in Section ~\ref{sec:result}:

\emph{\textbf{Exception Type Analysis.}} To support a deeper investigation of the 
stack traces every exception defined on a stack trace need to be classified according to its type
(e.g. Error, Exception or RuntimeException). The module responsible for this analysis 
uses Design Wizard~\cite{Brunet09} framework to inspect the bytecode of all exceptions reported on issues.
The nidyke walks up the type hierarchy of a Java exception until it reaches a base exception type.
This module analyzed the exception hierarchy of all exceptions defined on:
Android Platform (API level 19) - which includes all basic Java exceptions that 
can be thrown during the app execution. The exceptions thrown inside the apps
were classified using manual inspections as well as the exceptions thrown only
on previous versions of the Android platform.

\emph{\textbf{Exception Signaler Analysis.}}
This module is responsible for classifying each signaler according 
to its origin (i.e. Android Application Framework, Android Libcore, Application, Library). 
Table~\ref{tab:signalers} presents the heuristic adopted in this classification.
To enable such classification, we needed to feed this module the information
comprising all Java packages that compose: the Android Plaform;
 the Android Libcore; and each analyzed Application. 
To discover the packages composing the first two origins
we could reffer to the Android Specification.
To discover the packages composing each application, this module 
extracted the manifest files of each Android app
 - which defines the main packages that compose the applications.
 Whenever this file was not available, we recursevely analyzed the 
structure of source-code directories composing the app
- filtering out the cases on which the application 
also included the source-code of reused libraries.
Then based on this information and using pattern matching 
between the signaler name and the packages this module identifies 
the origin of the exception signalers. The exceptions were considered
 from libraries if their packages were not defined 
on Android Platform, core libraries nor on applications.
Table ~\ref{signalers} summarizes the signaler classification.

@@

\subsection{Manual Inspections}
In the context of the ExceptionMiner tool, manual inspections were used:
to help on the identification of packages composing the Android platform, 
libs and apps analyzed in this study (as described previously); and to 
identify the type of some exceptions reported on issues, 
that could not be automatically identified by the ExceptionMiner tool.
Some exceptions could not be automaticaly identified because they were 
not defined on the last version of the application (the one analyzed by ExceptionMiner),
or were defined on libraries not found among application resources.
In such cases we manually inspect the sourcecode or javadoc
 documentation available online of each of them. 
When the exception was not found on the previous versions, the exception where classified 
in our study as "Undefined".  Only 31 exceptions 
remained undefined - such exceptions were reported on 61 exception stack traces as 
presented in Table ~\ref{tab:typeroottab}.

\subsection{Replication Package}
All the data used in this study will be publicly available at https://github.com/souzacoelho/exceptionminer
Specifically we provide: (i) all issues related to Android projects found
in GITHub and Google Code used in this study; (iii) all stack traces extracted
from issues (ii) the results of manual inspection steps; (iv) the tool developed 
in this work to support stack trace extraction and distilling.

@@

This section presents the results of the study, providing both a
quantitative and qualitative analysis of the outcomes. The structure of this section
 is centered on questions that guied the search for fragilities (on the exception-related code)
based on exception stack trace information.

@@

After distilling the information available on the exception stack traces we could find 
the the exceptions commonly reported as the root causes of stack traces.
Table \ref{tab:topten} presents a list the top 10 root exceptions found in the study - 
 ranked by the number of distinct projects on which they were reported. 
The table shows how many times the signaler of such exception was a method defined on
the Android Platform, the Android libcore, the application itself or a third-party library -
 following the classification as presented in Table  \ref{tab:signalers}.

@@

We can observe that most of the exceptions on this list are implicitly thrown by the
 runtime environment due to programming mistakes  (e.g., out-of-bounds array index, division-by-zero, access to a null reference)
 or resource limitation (e.g., OutOfMemoryError). 
From this set the java.lang.NullPointerException was the most reported root cause (27,71\% ). 
If we consider the frequencly of NullPointerException 
across projects, we can observe that 51,96\% of all projects reported at least one exception stack 
trace on which the NullPoiterException was the root cause.

@@

As mentioned before, using the ExceptionMiner tool in combination with manual inspections we could identify
the root exception types (i.e., RuntimeException, Error, checked Exception) as well as its origins - which were 
identified based on the package names of the signalers related to it on the stack traces (see Section ~\ref{sec:exceptionminer}).
Table~\ref{tab:typeroottab} presents the types and origins of root exceptions of all stack traces analyzed.

>>>>

Until recent years, the availability of data, reporting such failures, was scarce. 
It comprised only few large scale projects (e.g. the Mozilla crash report dataset\footnote{https://code.google.com/p/promisedata/issues/detail?id=88}).
Recently we have observed a rapid increase in the number of open
source projects that make their version control systems as well as the issue tracking systems
publicly available in repository hosting sites such as GitHub and Google Code.
Developers and users of open-source projects uses such provided
 infrastrucute to share information (e.g., exception stack traces~\cite{bettenburg2008makes,schroter2010stack}) across 
the software development.

@@

This study performs a post mortem analysis of the exception stack traces embedded on issues
reported on Android projects hosted in GitHub and Google Code. 
The goal of this study was to investigate whether the reported exception stack traces
 can reveal common fragilities on the exception-related code.  
We call ``fragilities" characteristics on the exception-related code 
that favor the introduction of failures such as uncaught 
exceptions -  contributing to decrease the applicaiton robustness.
To guide this investigation we compiled general guidelines on how to use
exceptions proposed by Gosling~\cite{gosling2000java},
Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective}.
Then, using a custom tool called ExceptionMiner,
 which we develop specifically for this study, we mine stack traces from the issues reported 
on 482 Android projects hosted in GitHub and 157 projects hosted in Google Code.
Overall 159,048 issues were analyzed and 6,005 stack traces were extracted from them.
The exception stack trace analysis was augmented with additional information
extracted from JDK, Android Plaform and the applications using byte code 
and source code analysis. Manual inspection was used to leverage the understanding 
of exception stack traces and support further discussions and insights.

@@

The Android Application Framework supports an event-driven development model. 
Among the basic elements that comprise an Android application, we can cite the
 Activity class. The Activity is the main class of an Android app (there is usualy one
Activity per application window). Each Activity defines a set of hook methods  
(e.g., OnCreate(), onPause()) that are dispatched automatically by the Android 
framework (by inversion of control), in response to infrastructure and application events.
The Android Application Framework also provides several supporting APIs, such as 
the APIs to support concurrent development (e.g. AsyncTask) - 
which are usually needed due to the high demand for responsiveness 
in mobile environment (mobile apps can easily become 
unresponsive due to resource limitation and excessive network access).

@@

By convention an Error represents an unrecoverable condition which usually results
from failures detected by the Java Virtual Machine, such as OutOfMemoryError and
normally should not be handled inside the application. The user-defined exceptions 
can be either checked or runtime. The Java specification~\cite{gosling2000java} 
suggests that user-defined exceptions should be checked, because doing so 
the callers of a method will know about the exceptions that a method can throw, and then 
they can decide what to do about them. However, runtime exceptions can also be 
used to represent user-defined exceptional conditions.There is a long-lasting debate 
about the pros and cons of both approaches~\cite{javatut,stackoverlow,debate}
%\footnote{152 questions in Stackoverlow are related to this debate}.

@@

@@

@@

\emph{IV- All exceptions explicitly thrown by a library/framework reusable code should be documented.}
(\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}).
The exceptions thrown by a method are an important part of methods interface,
and is required to use the method properly. The checked exceptions are already
part of the  methods signature, and the method caller is aware of the checked
exceptions being thrown by it. According to~\cite{bloch2008effective}, it is
also wise to document the explicitly thrown runtime exceptions\footnote{This
excludes the implicitly signaled runtime exceptions related to programming
mistakes} as carefully as checked exceptions, especially when developing
 library/framework reusable code. Doing so, the clients of such library/framework
will be aware of all the exceptions a method can throw and can design the code
to deal with them ~\cite{Robil00}.

If the developer fails to follow such practices (specially when developing 
library/framework code) it will be difficult or even impossible for the caller to 
make effective use of such method~\cite{wirfs2006toward, bloch2008effective},
and protect the system against unforeseen exceptions thrown by the utility code,
which can lead to the uncaught exceptions - one of the main causes of 
systems crashes. These best practices guided our exploratory study described next.

@@

The goal of this exploratory mining study was to investigate whether the exception stack traces 
embedded on issues of Android projects (hosted in GitHub and Google Code), can reveal 
\emph{common fragilities} on the exception code of both the applications and the underlying  framework. 
Our study focused on open-source apps, since the information needed to perform our study
cannot be retrieved from commercial apps, whose issue report systems and 
source codes are not publicly available. The Android open-source apps were also 
target of other research studies ~\cite{Linar13,ahimed}.

@@

To support this investigation, we developed a tool called ExceptionMiner (Section~\ref{sec:exceptionminer})
which extracts the exception stack traces embedded on issues, 
and combines stack trace information with source code and bytecode 
analysis. Such combination was needed to enable the identification of 
fragilities on the exception-related code. Moreover, manual inspections were also used to leverage
 the understanding of stack traces and support further discussions and insights.
In this study we explore the domain quantitatively and highlight interesting cases by 
exploring cases qualitatively.  Figure~ref{fig:overview} gives an overview of the study, whose tasks are detailed next.

@@

This study used the dataset provided by the GHTorrent project~\cite{Gousi13}, 
an off-line mirror of the data  offered through the Github API.  
To identify Android projects, we performed a case insensitive search for the
term \textsf{android} in the repository's names and short descriptions.  
Up to 23 Feb 2014,  when we queried GHTorrent, such heuristic could
 find 2.542 repositories. Running the ExceptionMiner tool 
 in this set we observed that 589 projects had at least one issue containing a stack trace.

Then we performed further cleanup, inspecting the site of every Android
reporting at least one stack trace, to make sure that they represented real
mobile apps. During this clean up 106 apps were removed because they were either
example projects (i.e., toy projects) or tools to support Android development
(e.g. Selendroid, Roboeletric - tools to support the testing of Android apps).
The filtered set consisted of 482 apps. This set of 482 projects contained overall 31,592 issues from which 4,042 exception stack traces 
were extracted. Issues on Github are different from issues on dedicated bug tracking tools such as 
Bugzilla and Jira. The most important difference is that there are no predefined fields
  (e.g. severity and priority). Instead, Github uses a more open ended tagging system, on which
repositories are offered a pre-defined set of labels, but repository owners can modify 
them at will. Therefore, an issue may have none or an arbitrary set of labels depending 
on which repository it was created. Table~\ref{tab:lables} illustrates the ocurrences of different labels 
on the issues including exception stack traces.

@@

\begin{table}
  \centering
  \begin{tabular}{lr|lr}
    \hline
     \multicolumn{2}{c}{\bfseries{GitHub}} &  \multicolumn{2}{c}{\bfseries{Google Code}} \\
      \bfseries{Lable} &  \bfseries{\% Occurrences} &  \bfseries{Lable} &  \bfseries{\% Occurrences} \\
    \hline
empty &	54,24\% & Defect &	91,96\% \\
Defect &	39,56\%  & Enhancement  &	3,16\% \\
Enhancement &	0,57\% & Task	& 1,37\% \\
Support &	0,52\% & empty &	1,12\% \\
Problem &	0,36\% & StackTrace &	0,70\% \\
Others &	4,74\% &  Others &	1,68\% \\   
  \hline
  \end{tabular}
  \caption{Labels on issues including exception stack traces.}
  \label{tab:lables}
\end{table}

Based on the assumption that regardless the issue type, every exception stack
trace contains relevant information concerning the exception structure of the
projects analyzed - and therefore can review fragilities on the exception-related code -  
we opted for not restricting the analysis only on defect
issues\footnote{We conducted the same analysis on the defect issues and the top
exceptions are similar to the ones mentioned on defect issues. Due to space
limitation we limit to present the general analysis here. More detailed analysis
on the defect issues can be found at:
\url{https://github.com/souzacoelho/exceptionminer}}.
 Moreover, issues and pull requests are dual on Github; all pull requests have a corresponding 
``backing'' issue (i.e., an issue that is automatically generated per pull request). Therefore, we excluded such automatically generated
issues from our analysis.

\subsubsection{Android Apps in Google Code}
Google Code contains largely used open-source Android apps (e.g. K9Mail \footnote{K9Mail moved to Github but as a way of not loosing the project history 
it advises their users to report bugs on Google Code issue tracker.}), however, differently from GitHub, Google Code does not provide an API to access the information related
 to hosted projects \footnote{Google code used to provided a Web service to such repository, but it was deactivated in June 2013 what Google called a "clean-up action".}.
To overcome this limitation we needed to implement a Web Crawler (incorporated in ExceptionMiner tool described next) that navigated 
through the web interface of Google Code projects extracting all issues and issue comments and storing in a relational database for later analysis.

To identify Android projects in Google Code, we performed a similar heuristic: we performed a case insensitive search 
(on Google Code search interface) for the term ``android". On January 2014, when we queried Google Code, such heuristic could
 returned a list of 788  projects. This list comprised the seeds sent to our Crawler .

The Crawler retrieved all issues and comments for these projects.
From this set, 724 defined at least 1 issue. Running the ExceptionMiner tool 
 in this set we observed that 183 projects had at least one issue containing an exception stack trace.
 Then we performed further cleanup (similar to the one described previously) inspecting the site 
of each project. As a result we could identify 157 Android projects.  This set contained overall 127,456 issues,
 from which 1,963 exception stack traces were extracted. Table~\ref{tab:lables} illustrates the ocurrences of different labels 
on the issues including exception stack traces. Differently from GitHub, in Google Code most of 
the issues were labled as ``Defect". However, based on the same assumption described for the GitHub repository
 we considered all issues reporting stacks (regardless its labels).

@@

The ExceptionMiner is a mining tool which can connect to different issue repositories, 
extract issues, mine stack traces from them, distill exception stack trace information,
and enable the execution of different analyses by combining exception stack 
trace information with bytecode and sourcecode analysis. The main components of ExceptionMiner are as follows:

\emph{\textbf{Repository Connectors.}} This component enables the connection 
with issue repositories. In this study two main connectors were created: one which connects to 
GHTorrent database, and Google Code connector which is comprised of a Web Crawler,
 that can traverse Google Code Web interface and extract the project issues. 
Projects meta-data and the issues associated to each project are stored on a relational 
database.

\emph{\textbf{Exception Stack Trace Distiller.}}
This component distills the information that composes a stack trace.
 Some of the attributes extracted from the stack trace are:
 the root exception and its signaler, the exception wrappers and their corresponding signalers. 
This component also distills fine grained information of each attribute such as: the classes and packages associated to them.
This component is based on a combination between a regular expression based parser 
with heuristics able to identify and filter exception names and stack traces inline with text. 
In contrast to existing issue parsing solutions such as Infozilla, the parser
created in this work can discover stack traces inlined with logs files \footnote{In several 
exception stack traces, the exception frames were preceeded by logging information e.g., 
"03-01 15:55:01.609 (7924): at android.app.ActivityThread.access\$600(ActivityThread.java:127) 
"which could not be detected by existing tools.)}.

@@

As mentioned before, the tool enables a set of exception stack traces analyses by combining exception stack 
trace information with bytecode and sourcecode analysis. Each analysis is implemented as a different module 
on ExceptionMiner. Next, we present core analyses implemente on ExceptionMiner. Such analyses 
provide additional information to be used on further analyses described in Section~\ref{sec:result}:

\emph{\textbf{Exception Type Analysis.}} To support a deeper investigation of the 
stack traces every exception defined on a stack trace need to be classified according to its type
(e.g. Error, Exception or RuntimeException). The module responsible for this analysis 
uses Design Wizard~\cite{Brunet09} framework to inspect the bytecode of all exceptions reported on issues,
and a specific implementation to analise the source code of Java Exception classes.
It walks up the type hierarchy of a Java exception until it reaches a base exception type.
This module analyzed all exceptions defined on Android Platform (API level 19) - 
which includes all basic Java exceptions that can be thrown during the app execution,
and exceptions thrown by Android core libraries. 
Exceptions thrown inside the apps and libraries were classified through manual inspections 
as well as the exceptions thrown only on previous versions of the Android platform
(only the exceptions reported on stack traces needed to be inspected).

\emph{\textbf{Exception Signaler Analysis.}}
This module is responsible for classifying each signaler according 
to its origin (i.e. Android Application Framework, Android Libcore, Application, Library). 
Table~\ref{tab:signalers} presents the heuristic adopted in this classification.
To enable such classification, we needed to feed this module the information
comprising all Java packages that compose: the Android Plaform;
 the Android Libcore; and each analyzed Application. 
To discover the packages composing the first two origins
we could refer to the Android Specification.
To discover the packages composing each application, this module 
extracted the manifest files of each Android app
 - which defines the main packages that compose the applications.
 Whenever this file was not available, the tool recursively analyzed the 
structure of source-code directories composing the app
- filtering out the cases on which the application 
also included the source-code of reused libraries.
Then based on this information and using pattern matching 
between the signaler name and the packages this module identifies 
the origin of the exception signalers. The exceptions were considered
 from libraries if their packages were not defined 
on Android Platform, or on core libraries or on applications.
Table~\ref{tab:signalers} summarizes the signaler classification.

@@

\subsection{Manual Inspections}
In the context of the ExceptionMiner tool, manual inspections were used to:
(i) support on the identification of packages composing the Android platform, 
libs and apps analyzed in this study (as described previously); and (ii)  
identify the type of some exceptions reported on issues, 
that were not be automatically identified by the ExceptionMiner tool.
When the exception was not found automatically, nor manually (because they were related an app or lib the previous versions
that were no more availble), the exception were classified 
in our study as "Undefined".  Only 31 exceptions 
remained undefined - such exceptions were reported on 61 exception stack traces as 
presented in Table ~\ref{tab:typeroottab}.

\subsection{Replication Package}
All the data used in this study is publicly available at https://github.com/souzacoelho/exceptionminer
Specifically we provide: (i) all issues related to Android projects found
in GITHub and Google Code used in this study; (iii) all stack traces extracted
from issues (ii) the results of manual inspection steps; (iv) the tool developed 
in this work to support stack trace extraction and distilling.

@@

This section presents the results of the study, providing both a
quantitative and qualitative analysis of the outcomes. The structure of this section
 is centered on questions that guided the search for fragilities (on the exception-related code)
based on exception stack trace information.

@@

After distilling the information available on the exception stack traces we could find 
the the exceptions commonly reported as the root causes of stack traces.
Table \ref{tab:topten} presents a list of the top 10 root exceptions found in the study - 
 ranked by the number of distinct projects on which they were reported. 
This table also shows how many times the signaler of such exception was a method defined on
the Android Platform, the Android libcore, the application itself or a third-party library -
 following the classification as presented in Table  \ref{tab:signalers}.

@@

We can observe that most of the exceptions on this list are implicitly thrown by the
 runtime environment due to programming mistakes  (e.g., out-of-bounds array index, division-by-zero, access to a null reference)
 or resource limitation (e.g., OutOfMemoryError). 
From this set the java.lang.NullPointerException was the most reported root cause (27,71\%). 
If we consider the frequency of NullPointerException 
across projects, we can observe that 51,96\% of all projects reported at least one exception stack 
trace on which the NullPoiterException was the root cause.

@@

As mentioned before, using the ExceptionMiner tool in combination with manual inspections we could identify
the root exception types (i.e., RuntimeException, Error, checked Exception) as well as its origins - which were 
identified based on the package names of the signalers related to it on the stack traces (Section~\ref{sec:exceptionminer}).
Table~\ref{tab:typeroottab} presents the types and origins of root exceptions of all stack traces analyzed.
